Simulating the Linux page cache in distributed applications with the WRENCH library

### 2020-12-08 paper_io_sim.tex
replaced tabs with spaces (again)
Edits in method

        In this paper, we present WRENCH-Cache, a simulation model of I/O with page
        cache incorporated into WRENCH, a workflow management system simulator based of
        the popular SimGrid framework. The goal of WRENCH-Cache is to provide users
        with an accurate representation of a workflow execution, when file system page
        cache is used. WRENCH-Cache is able to simulate both writeback and writethrough
        cache which would normally be available to local and NFS-based filesystems.

@@

    \section{Method}
    \label{method}
    In this section, we present our page cache simulation model for file I/O,
    flusing, and eviction. We detail the design and implementation of two simulators:
    (1) a standalone Python prototype to evaluate the simulation model independently of any other 
    tool, and (2) WRENCH-cache, an extension of WRENCH with our page cache model.
    We also describe experimental scenarios to evaluate our simulators in various configurations.

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IOController. The IOController interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.
    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the simulated writeback page cache.
           Applications send file read or write requests to the
           IOController that orchestrates flushing, eviction, cache
           and disk accesses with the MemoryManager. Storage
           devices (memory and disk) simulate file transfers with
           bandwidth sharing.}
           \label{fig:interaction}
    \end{figure}

@@

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cache I/Os synchronously, whereas
    the second one, operating in the background, periodically searches for
    expired dirty data in page cache LRU lists and flushes them to disk. We
    use existing storage simulation models~\cite{lebre2015} to simulate
    memory, characterized by its storage capacity, read and write
    bandwidths, and latency. Reusing existing storage models allows us to
    simulate bandwidth sharing between concurrent memory accesses.

@@

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
    % Modeling page cache LRU lists as lists of data blocks reduces the
    % overhead of the simulator, as data blocks are obviously less
    % numerous than file pages. At the same time, it achieves the accuracy in 
    % I/O time simulation that can be seen in the later section.
    % From Tristan: I don't think the sentences above are necessary,
    % they are quite repetitive.

@@

    \subsubsection{Reads and Writes}

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and put at the top of the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and put it in the active list. 
    If the blocks are dirty, we move them to the active list to preserve their entry time. 
    Because the chunk size and block size can be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list.}
           \label{fig:read_order}
    \end{figure}

    For file writes, we assume that all data to be written is 
    uncached. Thus, each time a chunk is written, we create a block of dirty data 
    and add it to the inactive list.

@@

    The main simulated thread in the Memory Manager can flush the
    cache to write dirty data to disk. The data flushing simulation
    function takes the amount of data to flush as parameter. While
    the amount of data to flush is not reached and there are dirty
    blocks remaining in cache, this function traverses the sorted
    inactive list, then the sorted active list, and writes the
    least recently used dirty block to disk having set its dirty
    flag to 0. In case the amount of data to flush requires that a
    block be partially flushed, the block is split in two blocks,
    one that is flushed and one that remains dirty. Flushing time
    to disk is simulated with the storage model.

    Similarly to data flushing, our cache eviction simulation runs in
    the main thread. It frees up the page cache by traversing and deleting 
    least recently used clean data blocks in the inactive list.
    The amount of data to evict is passed as a parameter and data blocks are deleted 
    from the inactive list until total evicted data reaches the required amount,
    or until there is no clean block left in the list.
    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
    and only one of them is evicted.
    The cache eviction simulation does not contribute to the simulated time 
    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.

@@

    Periodical flushing is simulated in the Memory Manager
    background thread. As in the Linux kernel, a dirty block
    in our model is considered expired if (1) it is dirty, and (2)
    the duration since its entry time is longer than a
    predefined expiration time. 
    Periodical flushing is simulated as an infinite loop in which 
    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
    % From Tristan: the algorithm is quite straightforward, I don't think this is 
    % necessary
    % In each repetition, Memory Manager finds expired dirty blocks in two 
    % page cache LRU lists (line 8), simulates writes of data of these blocks 
    % to disk (line 9), and mark them as clean (line 10).
    % If the flushing time does not exceed our time interval, the thread is put 
    % to sleep for the remaining time (lines 11-13). 
    % Then based on the state of the host at current simulated time, 
    % the algorithm continues or finishes the loop.
    Because periodical flushing is simulated as a background thread, it can happen concurrently
    with disk I/O initiated by the main thread. This parallelism is handled by the 
    storage model and reflected in simulated I/O time.

@@

    \begin{algorithm}\caption{File chunk read simulation of IOController}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn))
               \State cache\_read = cs - disk\_read
            \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
            \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                \State sm.read(disk\_read)  
                \State mm.add\_to\_cache(disk\_read, fn)     
            \EndIf
            \If {cache\_read $>$ 0} \Comment{Read cached}
                \State mm.cache\_read(cache\_read)  
            \EndIf
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IOController which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache \tristan{Create a variable
    "required\_mem" in the algorithm to clarify}.
    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). When called with negative arguments, functions
    flush and evict simply return and do not do anything. Then, if there is
    uncached data, a disk read is simulated with disk model (line 12)
    \tristan{this is not consistent with the new Figure 1 where disk reads
    are done by the mm}, and this data of the file is added to cache (line
    13).
    Finally, if cached data needs to be read, the Memory Manager is called
    to simulate a cache read  
    and update the corresponding data blocks accordingly (line 16).

    \begin{algorithm}\caption{File chunk write simulation of IOController}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write(fn, mem\_amt) 
            \EndIf
            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)  
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IOController is not supposed to interact directly with sm}
            \EndIf

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
    the IOController. 
    % Data is written to cache until the amount of dirty data in the cache 
    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
    Our algorithm initially checks the  amount of data that 
    can be written as dirty data given the dirty ratio (line 6).
    If this amount is greater than 0, the Memory Manager is requested to evict 
    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained somewhere}.
    After eviction, the amount of data that can be written to 
    page cache is calculated (line 9), and a cache write is simulated (line 10).
    To write the remaining data, we flush and 
    evict data as possible so that this data can be 
    written to cache (line 13-14). 
    The Memory Manager does not guarantee that the amount of flushed 
    and evicted data is enough to write the remaining data. 
    In the case it is less than the required amount, the remaining data 
    being written to cache can be flushed and evicted right away. 
    Thus, the amount of remaining data written and kept in cache 
    after the write is limited by the amount free memory, and this amount 
    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
    Finally, the disk write is simulated with the remaining data amount (line 16), 
    and the chunk write algorithm finishes \tristan{see comment in algorithm}.

@@

            We also implemented our model in WRENCH, separately from existing WRENCH components. 
            Page cache simulation is available through a user API and can be activated using a command-line argument.
            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
            \tristan{which commit did you use?}

@@

        Our experiments compare real executions with the original WRENCH,
        with our WRENCH-cache extension, and with our prototype Python simulator. We
        evaluate our page cache simulation in single-threaded and
        multi-threaded applications, accessing data on local and remote
        file systems.

        We simulate an application with three single-core, sequential tasks
        where each task reads the file produced by the previous task,
        increments every byte of this file to emulate real processing, and
        writes the resulting data to disk. We also use a real application
        \textcolor{red}{[Maybe briefly describe this application's purpose
        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
        applicability of our simulation model.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since SimGrid only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|c
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network bandwidth & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

@@

            Our first experiment simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk.

@@

            Our third experiment occurs in the same configuration as the
            previous one, albeit on a 400-GB NFS-mounted remote disk
            mounted from another compute node. \tristan{As is commonly
            configured in HPC environments to avoid data loss,?} NFS client
            and server read caches are enabled, but server writes are configured with
            write-through cache instead of write back \tristan{how about client write
            cache?}. Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding them to the server
            cache \tristan{it it part of the IOController? it would deserve
            a few sentences in the methods.}.

>>>>

        In this paper, we present WRENCH-cache, a simulation model of I/O with page
        cache incorporated into WRENCH, a workflow management system simulator based of
        the popular SimGrid framework. The goal of WRENCH-cache is to provide users
        with an accurate representation of a workflow execution, when file system page
        cache is used. WRENCH-cache is able to simulate both writeback and writethrough
        cache which would normally be available to local and NFS-based filesystems.

@@

    \section{Method}
    \label{method}
    In this section, we present our page cache simulation model for file I/O,
    flusing, and eviction. We implement this model in two simulators:
    (1) a standalone Python prototype, and (2) WRENCH-cache, an extension of the WRENCH library.
    We also describe experimental scenarios to evaluate our simulators in various configurations.

    We separate our simulation model in two components, the IO
    Controller and the Memory Manager, which together simulate 
    file reads and writes in Linux (Figure~\ref{fig:interaction}). 
    To read or write a file chunk, a simulated application sends a
    request to the IO Controller. The IO Controller interacts as needed with
    the Memory Manager to free memory through flushing or eviction,
    and to read or write to cache. The Memory Manager
    implements these operations, simulates periodical flushing
    and eviction from cache, and reads or writes to disk when necessary.
    \begin{figure}
           \centering
           \includegraphics[width=0.85\columnwidth]{figures/interaction.pdf}
           \caption{Overview of the simulated writeback page cache \tristan{does it simulate only writeback? See comment in NFS section.}.
           Applications send file read or write requests to the
           IO Controller that orchestrates flushing, eviction, cache
           and disk accesses with the Memory Manager. Concurrent accesses to storage
           devices (memory and disk) are simulated using existing models.}
           \label{fig:interaction}
    \end{figure}

@@

    The Memory Manager simulates two parallel threads: the main one
    implements flushing, eviction, and cache I/Os synchronously, whereas
    the second one, operating in the background, periodically searches for
    expired dirty data in page cache LRU lists and flushes them to disk. We
    use existing storage simulation models~\cite{lebre2015} to simulate disk and
    memory, characterized by its storage capacity, read and write
    bandwidths, and latency. Reusing existing storage models allows us to
    simulate bandwidth sharing between concurrent memory accesses and concurrent disk accesses.

@@

    We model page cache LRU lists as 
    two lists of data blocks, an active list and an inactive list, both ordered by 
    last access time (earliest first, see Figure~\ref{fig:lrulist}).
    As in the kernel, our simulator limits the size of the active list to
    twice the size of the inactive list, by moving least recently 
    used data blocks from the active list to the inactive list~\cite{gorman2004understanding, linuxdev3rd2010}.
    % Modeling page cache LRU lists as lists of data blocks reduces the
    % overhead of the simulator, as data blocks are obviously less
    % numerous than file pages. At the same time, it achieves the accuracy in 
    % I/O time simulation that can be seen in the later section.
    % From Tristan: I don't think the sentences above are necessary,
    % they are quite repetitive.

@@

    \subsubsection{Reads and writes}

    Our simulation model supports chunk-by-chunk file accesses
    with a user-defined chunk size. However, for simplicity, we assume that file pages are 
    accessed in a round-robin fashion rather than fully randomly. 
    Therefore, when a file is read, cached data is read only after all uncached data was read, and data from the inactive list is read
    before data from the active list
    (data reads occur from left to right in Figure~\ref{fig:read_order}).
    When a chunk of \emph{uncached} data is read, a new clean block is created 
    and appended to the inactive list. 
    When a chunk of \emph{cached} data is read, one or more existing data blocks in the LRU lists are accessed.
    If these blocks are clean, we merge them together, update the access time and size of the resulting block, 
    and append it to the active list. 
    If the blocks are dirty, we move them independently to the active list, to preserve their entry time. 
    Because the chunk and block sizes may be different, there are situations
    where a block is not entirely read. 
    In this case, the block is split in two smaller blocks and one of them is re-accessed.
    % From Tristan: not sure where to put this nor if it's necessary:
    % , in which chunks are read/written
    % until file is entirely read/written.
    \begin{figure}
           \centering
           \includegraphics[width=\columnwidth]{figures/read_order.pdf}
           \caption{File data read order. Data is read from left to right: uncached data 
           is read first, followed by data from the inactive list, and finally data from the active list. \tristan{Colors 
           shouldn't match that in Fig 1 as they dont represent the same thing.}}
           \label{fig:read_order}
    \end{figure}

    For file writes, we assume that all data to be written is 
    uncached. Thus, each time a chunk is written, we create a block of dirty data 
    and append it to the inactive list.

@@

    The main simulated thread in the Memory Manager can flush or evict data from the
    memory cache. The data flushing simulation
    function takes the amount of data to flush as parameter. While
    the amount of data to flush is not reached and there are dirty
    blocks remaining in cache, this function traverses the sorted
    inactive list, then the sorted active list, and writes the
    least recently used dirty block to disk, having set its dirty
    flag to 0. In case the amount of data to flush requires that a
    block be partially flushed, the block is split in two blocks,
    one that is flushed and one that remains dirty. Flushing time
    to disk is simulated by the storage model.

    The cache eviction simulation also runs in
    the main thread. It frees up the page cache by traversing and deleting 
    least recently used clean data blocks in the inactive list.
    The amount of data to evict is passed as a parameter and data blocks are deleted 
    from the inactive list until the evicted data reaches the required amount,
    or until there is no clean block left in the list.
    If the last evicted block does not have to be entirely evicted, the block is split in two blocks,
    and only one of them is evicted.
    The cache eviction simulation does not contribute to the simulated time 
    since cache eviction time is negligible in real systems \tristan{a reference would be welcome}.

@@

    Periodical flushing is simulated in the Memory Manager
    background thread. As in the Linux kernel, a dirty block
    in our model is considered expired if (1) it is dirty, and (2)
    the duration since its entry time is longer than a
    predefined expiration time. 
    Periodical flushing is simulated as an infinite loop in which 
    the Memory Manager searches and flushes these expired dirty blocks to disk (Algorithm~\ref{alg:pdflush}). 
    % From Tristan: the algorithm is quite straightforward, I don't think this is 
    % necessary
    % In each repetition, Memory Manager finds expired dirty blocks in two 
    % page cache LRU lists (line 8), simulates writes of data of these blocks 
    % to disk (line 9), and mark them as clean (line 10).
    % If the flushing time does not exceed our time interval, the thread is put 
    % to sleep for the remaining time (lines 11-13). 
    % Then based on the state of the host at current simulated time, 
    % the algorithm continues or finishes the loop.
    Because periodical flushing is simulated as a background thread, it can happen concurrently
    with disk I/O initiated by the main thread. This is taken into account by the 
    storage model and reflected in simulated I/O time.

@@

    \begin{algorithm}\caption{File chunk read simulation of IO Controller}
    \label{alg:read}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{fs}{file size (assumed to fit in memory)}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
               \State disk\_read = min(cs, fs - mm.cached(fn)) \Comment{To be read from disk} 
               \State cache\_read = cs - disk\_read \Comment{To be read from cache} 
            \State mm.flush(cs + disk\_read - mm.free\_mem - mm.evictable, fn) 
            \State mm.evict(cs + disk\_read - mm.free\_mem, fn) 
            \If {disk\_read $>$ 0}  \Comment{Read uncached data}    
                \State sm.read(disk\_read)  
                \State mm.add\_to\_cache(disk\_read, fn)     
            \EndIf
            \If {cache\_read $>$ 0} \Comment{Read cached}
                \State mm.cache\_read(cache\_read)  
            \EndIf
        \end{algorithmic}
    \end{algorithm}
    As mentioned previously, our model reads and writes file chunks in a
    round-robin fashion. To read a file chunk, simulated applications send
    chunk read requests to the IO Controller which processes them using
    Algorithm~\ref{alg:read}. First, we calculate the amount of uncached
    data that needs to be read from disk, and the remaining amount is read
    from cache (line 7-8). The amount of memory required to read the chunk
    is calculated, corresponding to a copy of the chunk in anonymous memory
    and a copy of the chunk in cache \tristan{Create a variable
    "required\_mem" in the algorithm to clarify}.
    \tristan{add a call to mm in the algorithm to increment anonymous memory. Mention in the experiments that anonymous memory is flushed after each task.}
    If there is not enough available memory, the Memory Manager is called
    to flush dirty data (line 9). If necessary, flushing is complemented by
    eviction (line 10). Note that, when called with negative arguments, functions
    \texttt{flush} and \texttt{evict} simply return and do not do anything. Then, if the block requires
    uncached data, a disk read is simulated with the disk model (line 12)
    \tristan{this is not consistent with the new Figure 1 where disk reads
    are done by the mm}, and the data read is added to cache (line
    13).
    Finally, if cached data needs to be read, the Memory Manager is called
    to simulate a cache read  
    and update the corresponding data blocks accordingly (line 16).

    \begin{algorithm}\caption{File chunk write simulation of IO Controller}
    \label{alg:write}
        \small
        \begin{algorithmic}[1]
            \Input
                \Desc{cs}{chunk size}
                \Desc{fn}{file name}
                \Desc{mm}{MemoryManager object}
                \Desc{sm}{storage simulation model}
               \EndInput
            \State remain\_dirty = dirty\_ratio * mm.avail\_mem - mm.dirty
            \If {remain\_dirty $>$ 0} \Comment{Write with memory bandwidth}
                \State mm.evict(min(cs, remain\_dirty) - mm.free\_mem)
                \State mem\_amt = min(cs, mm.free\_mem)
                \State mm.write(fn, mem\_amt) 
            \EndIf
            \If {mem\_amt $<$ cs}  \Comment{Write with disk bandwidth}
                \State mm.flush(cs - mem\_amt)  
                \State mm.evict(cs - mem\_amt  - mm.free\_mem) 
                \State mm.add\_to\_cache(fn, min(cs - mem\_amt, mm.free\_mem)) \tristan{why is data written to cache with "addtocache" here and with "write" at line 10?}
                \State sm.write(cs - mem\_amt) \tristan{According to Fig 1 the IO Controller is not supposed to interact directly with sm}
            \EndIf

        \end{algorithmic}
    \end{algorithm}
    Algorithm~\ref{alg:write} describes our simulation of chunk writes in 
    the IO Controller. 
    % Data is written to cache until the amount of dirty data in the cache 
    % exceeds the dirty ratio times the amount of available memory. Data is then written to disk.
    Our algorithm initially checks the  amount of data that 
    can be written as dirty data given the dirty ratio (line 6).
    If this amount is greater than 0, the Memory Manager is requested to evict 
    data from cache if necessary (line 8) \tristan{the difference between available and free memory should be explained the first time it is referred}.
    After eviction, the amount of data that can be written to 
    page cache is calculated (line 9), and a cache write is simulated (line 10).
    To write the remaining data, we flush and 
    evict data as possible so that this data can be 
    written to cache (line 13-14). 
    The Memory Manager does not guarantee that the amount of flushed 
    and evicted data is enough to write the remaining data. 
    In the case it is less than the required amount, the remaining data 
    being written to cache can be flushed and evicted right away. 
    Thus, the amount of remaining data written and kept in cache 
    after the write is limited by the amount free memory, and this amount 
    is added to cache (line 15) \tristan{the previous two sentences are very unclear, I don't understand them so I can't reformulate.}. 
    Finally, the disk write is simulated with the remaining data amount (line 16), 
    and the chunk write algorithm finishes \tristan{see comment in algorithm}.

@@

            We also implemented our model in WRENCH, separately from existing WRENCH components. 
            We exposed page cache simulation through a user API, activated with a command-line argument.
            We used SimGrid 3.25 and WRENCH 1.6. SimGrid source code is available at \url{https://framagit.org/simgrid/simgrid},
            and WRENCH is at \url{https://github.com/wrench-project/wrench/tree/master/src/wrench/services/memory}.
            \tristan{which commit did you use?}

@@

        Our experiments compare real executions with our Python prototype,
        with the original WRENCH, and with our WRENCH-cache extension. We
        evaluate our page cache simulation in single-threaded and
        multi-threaded applications, accessing data on local and remote
        file systems.

        We simulate an application that consists of three single-core, sequential tasks
        where each task reads the file produced by the previous task,
        increments every byte of this file to emulate real processing, and
        writes the resulting data to disk. We also use a real application
        \textcolor{red}{[Maybe briefly describe this application's purpose
        (e.g. a standard fMRI preprocessing piepeline)]}, to evaluate the
        applicability of our simulation model.

@@

        To parametrize our simulators, we benchmarked the bandwidths of
        memory, local disk, remote disk (NFS), and network
        (Table~\ref{table:benchmark}). Since WRENCH currently only supports
        symmetrical bandwidths, we use the mean of the read and write
        bandwidth values in our experiments.
            \begin{table}[htbp]
            \centering
            \begin{tabularx}{\columnwidth}{|c
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X
            |>{\centering\arraybackslash}X|}
            \hline
                Device  & Cluster (real) & Python simulator & WRENCH simulator\\
            \hline
                Memory read  & 6860    & 4812     & 4812\\
                Memory write & 2764    & 4812 & 4812\\
                Local disk read & 510 & 465 & 465\\
                Local disk write & 420 & 465     & 465\\
                Remote disk read & 515 & - & 445\\
                Remote disk write & 375 & - & 445\\
                Network bandwidth & 3000 & - & 3000\\
            \hline
            \end{tabularx}
            \caption{Bandwidth benchmarks (MBps) and simulator configurations.
            The bandwidths used in the simulations are the average of the read and write bandwidths
            measured in the cluster.
            Remote disk and network accesses are not simulated in the Python simulator. \tristan{table could be prettified}}
            \label{table:benchmark}
            \end{table}

@@

            Our first experiment simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk. The goal is to validate the simulation model 
            in absence of any concurrent disk access.

@@

            Our third experiment occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB remote NFS disk
            mounted from another compute node. \tristan{As is commonly
            configured in HPC environments to avoid data loss,?} NFS client
            and server read caches are enabled, but server writes are configured with
            write-through cache instead of write back \tristan{how about client write
            cache?}. Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding them to the server
            cache \tristan{it it part of the IO Controller? it would deserve
            a few sentences in the methods.}.
